Numerical investigation of viscous flow instabilities through heterogeneous porous media. Advantages and limitations of unstructured and adaptive meshes in 2D and 3D domains under different mobility scenarios.

### 2017-08-11 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Updates on Section 4.5

\medskip
As mentioned in Section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of flow patterns, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any instabilities in the flow front. Figure \ref{fig:3DChannel_PermMesh} shows the initial permeability distribution for the simulations described in this section. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a uniform velovity of 2.5 cm.s$^{-1}$. No-slip conditions are applied everywhere else (\ie no flux across the boundaries) except at the right hand side of the domain where the fluid is allowed to leave the domain. Initially, the domain was partially saturated with with non-wetting fluid, \ie $S_{nw}(t=0) = 0.80$, which was displaced by a pure (wetting) fluid. The domain was designed to capture preferential flow pathways, as shown in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

\medskip
Figure \ref{fig:3DChannel_planes} gives a comprehensive description of the contours of phase $1$ Darcy velocity, in times related to the beginning and just before the end of simulation. While the saturation level for phase $1$, remains between the values of $0.25-0.75$ the main difference between the figures \ref{fig:3DChannel_planes} (a) and (b) is the intensity of the flow as this can be given by the darcy velocity contours. In the beginning of the simulation most of the contours for phase $1$ are populated and achieve their higher values closer to the inlet region. At this point a comparison between the figures \ref{fig:3DChannel_planes}(a)-\ref{fig:3DChannel_sat}(a) and \ref{fig:3DChannel_planes}(b)-\ref{fig:3DChannel_sat}(f) as they both describe the darcy velocity contours and the saturation front for the same timesteps, is necesserary.

\medskip
In sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}, a $2$D domain was used to describe one or more finger formations. These flow instabilities were triggerded from the permeability differences in the domain. Using the $3D$ domain, the difference between a flow instability and the preferential flow path that may include smaller fingers is more obvious. Figures \ref{fig:3DChannel_sat}(a)-(f) describe the flow path in various time-steps alongside with the progression of the saturation front that now becomes more obvious. Heterogeneity in the reservoir controls the spatial distribution of porosity ($\phi =0.2$ in the current case) and permeability. The permeability depends on the number, size and connectivity of the pores. An on purpose implementation of significant high permeability zones (compared with the lower ones) also know as higher permeability channels, allows the domain to give a good approximation of the so called "thief zones".

>>>>

\medskip
As mentioned in Section \ref{section:ViscousInstabilities}, \textit{viscous fingering} is characterized by the formation of flow patterns, called viscous fingers. The geometrical evolution of these patterns provides good indications on capturing and tracking any instabilities in the flow front. Figure \ref{fig:3DChannel_PermMesh} shows the initial permeability distribution for the simulations described in this section. The surface on the left hand side of the domain is the area where fluid $1$ is injected with a uniform velocity of 2.5 cm.s$^{-1}$. No-slip conditions are applied everywhere else (\ie no flux across the boundaries) except at the right hand side of the domain where the fluid is allowed to leave the domain. Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0) = 0.80$, which was displaced by a pure (wetting) fluid. The domain was designed to capture preferential flow pathways, as shown in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

\medskip
Flow pathways are clearly indicated in Fig.~\ref{fig:3DChannel_planes}, where fluid $1$ Darcy velocity vectors overlapped with fluid saturation are shown. Fluid displacement occurs mainly in regions of relatively larger permeability, leaving other parts of the domain (\ie regions with low permeability distribution) with little or no momentum, indicating that at these regions the non-wetting fluid (\ie fluid 2) are not displaced.

\medskip
In Sections \ref{section:results:setup}-\ref{section:results_hete_fix_adapt}, a $2$D domain was used to describe one or more finger formations. These flow instabilities were triggerded from the permeability differences in the domain. Using the $3D$ domain, the difference between a flow instability and the preferential flow path that may include smaller fingers is more obvious. Figures \ref{fig:3DChannel_sat}(a)-(f) describe the flow path in various time-steps alongside with the progression of the saturation front that now becomes more obvious. Heterogeneity in the reservoir controls the spatial distribution of porosity ($\phi =0.2$ in the current case) and permeability. The permeability depends on the number, size and connectivity of the pores. An on purpose implementation of significant high permeability zones (compared with the lower ones) also know as higher permeability channels, allows the domain to give a good approximation of the so called "thief zones".
